---
title: "Lab 6: GeoTIFF, Kriging"
author: "Jessica Jagdeo"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Attach packages:
```{r}

library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat) # variogram

```

## Grand Canyon GeoTIFF
```{r}

gc_dem <- raster(here("data", "gc_dem.tif"))

# Look at it using plot():
plot(gc_dem)

# Check the CRS:
gc_dem@crs

# Check the extent (bounding box):
gc_dem@extent 
# Gives us min and max in terms of meters

# Create WGS84 w/ latlong:
wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

# Reproject data into latlong:
gc_reproj <- projectRaster(gc_dem, crs = wgs84, method = "bilinear")

gc_reproj@extent
# Now, it has been reprojected into latlong degrees

```

### Crop raster to a smaller area (bounding box):
```{r}

bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), "SpatialPolygons")

```

### Make the CRS of our bounding box the same as for gc_reproj:
```{r}

crs(bounds) <- crs(gc_reproj)

```

###Now let's crop our original spatial data:
```{r}

gc_crop <- crop(gc_reproj, bounds)

plot(gc_crop)

```

### Resample using the aggregate() function:
```{r}

gc_agg <- raster::aggregate(gc_crop, fact = 1)

plot(gc_agg)

```

### Now let's get safe and sound in ggplot:

First, convert data to a data frame:
```{r}

gc_df <- as.data.frame(gc_agg, xy = TRUE) # Use xy = TRUE to include latlong coordinates

ggplot(data = gc_df, aes(x = x, y = y)) +
  geom_raster(aes(fill = layer)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c(
    "purple", 
    "magenta", 
    "orange", 
    "white"
  ))

```

### How can I just select cells that match given criteria?

Let's say I have some species of vegetation that will only grow when gc_dem is between 1000 and 1500
```{r}

gc_hab <- gc_crop

# Set any cells outside of 1000 - 1500 to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

# Plot with base plot:
plot(gc_hab)

```

Now, let's make this interactive with tmap:
```{r}

tmap_mode("view")

tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")

```

### Kriging rain in Kansas

Read in KS counties shapefile data:
```{r}

ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp"))

# Use base plot:
plot(ks_counties)

# Check CRS using SF:
st_crs(ks_counties)
# No coordinate reference system, so let's set it

# Set CRS to EPSG 4326:
st_crs(ks_counties) <- 4326

plot(ks_counties)

```

Practice in ggplot:
```{r}

ggplot(data = ks_counties) +
  geom_sf()

```

Now, let's read in the rainfall data:
```{r}

ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names()

```

Update the ks_rain data to be recognized at spatial points:
```{r}

ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"),
                  crs = 4326)

```

```{r}

ggplot() +
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf,
          aes(color = amt, size = amt)) # Makes a bubble plot, varying size and color

```

### Kriging to predict rainfall:
```{r}

ks_sp <- as_Spatial(ks_sf)
class(ks_sp) # Now it's SpatialPointsDataFrame

```

### Make a spatial pixels grid that we'll make predictions over:
```{r}

lat <- seq(37, 40, length.out = 200) # Create pixels evenly spaced in latitude

long <- seq(-94.6, -102, length.out = 200) # Create pixels evenly spaced in longitude

# Now, we're going to make this into a spatial grid:
grid <- expand.grid(lon = long, lat = lat)

grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326) # Converting to sf

grid_sp <- as_Spatial(grid_sf) # Converting to spatial

plot(grid_sp)

```

### Make a variogram
```{r}

ks_vgm <- variogram(amt ~ 1, data = ks_sp)

plot(ks_vgm)

# My estimates for the variogram parameters:
# Nugget = 0.2
# Sill = 0.8
# Range = 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))

plot(ks_vgm, ks_vgm_fit)

ks_vgm_fit

```

### Now, krige:
```{r}

ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

```

```{r}

spplot(ks_krige, "var1.pred")

```

### Make a data frame of kriged predictions:
```{r}

ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

# Converting this to sf object:
rain_sf <- st_as_sf(ks_df, coords = c("longitude", "latitude"), crs = 4326)

ggplot(rain_sf) + 
  geom_sf(aes(color = var1.pred))

```

### Crop to the actual KS outline:
```{r}

ks <- read_sf(dsn = here("data", "states"),
              layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)

```

```{r}

# Find the intersection of the two:

rain_sf_ks <- st_intersection(rain_sf, ks)

ggplot(data = rain_sf_ks) +
   geom_sf(aes(color = var1.pred)) +
  scale_color_gradientn(colors = c("white","yellow","magenta","purple")) +
  theme_minimal()

```

